Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-07-23 concepts.tex
Results data
missing references
AA: intro + related work reviewed

The {\em Pareto-optimal set} of $S$ for the application $b$ is the
set:
\[ \mathcal{P}(S,b) = \{ \mathbf{c} \in \mathcal{C}^*(S) \ : \nexists \ \mathbf{c}' \in \mathcal{C}^*(S), \mathbf{c}' \succ \mathbf{c} \} \]
that is, the set of configurations $\mathbf{c} \in
\mathcal{C}^*(S)$ not dominated by any other configuration.
Pareto-optimal configurations are configurations belonging to the
Pareto-optimal set and the {\em Pareto-optimal front} is the image
of the Pareto-optimal configurations, i.e. the set:
\[ \mathcal{P}_{F}(S,b) = \{ \mathbf{o} : \mathbf{o} = E(\mathbf{c},b), \mathbf{c} \in \mathcal{P}(S,b) \} \]

@@

\begin{figure}[h]
\includegraphics[width=0.9\columnwidth]{img/Pareto_set_and_front}

\caption{Relation between Pareto set (its ``parameter space representation'')
and Pareto front (``objective space representation'')}

\subsection{Required Definitions and Concepts}

@@

We also consider, in measuring how much iteresting a region is, the
``innovation'' that it adds. If, during a design space exploration,
a Pareto front is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable ``innovation'',
because the way it fulfils the objectives is very similar to other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is very distant from previous ones is remarkable,
because it let the experimenter realize that the objectives can be
fulfilled in a completely different way from what he already found
during exploration. So a region will be evaluated ineresting as much
as how different are its added Pareto front points from the ones already
found.
\begin{rem}
We will consider only parameters $p_{i}$ with ordering, i.e. such
that, if $P_{i}$ is a domain, $\forall a,b\in P_{i}$ it is possible
to say $a<b$ or $a=b$ or $a>b$.
\end{rem}
%TODO: DESCRIVERE COME SI PENSA DI TRATTARE GLI SCENARI IN CUI ESISTONO
%PARAMETRI SENZA ORDINAMENTO{]}
\begin{defn}
\emph{Parameter interval}

@@

\begin{defn}
\emph{\label{pers02.def:Contiguous-intervals}Contiguous intervals}

@@

>>>>

The {\em Pareto-optimal set} of $S$ for the application $b$ is the
set:
\begin{defn}
\label{pers02.def:Pareto-set}
\[ \mathcal{P}(S,b) = \{ \mathbf{c} \in \mathcal{C}^*(S) \ : \nexists \ \mathbf{c}' \in \mathcal{C}^*(S), \mathbf{c}' \succ \mathbf{c} \} \]
\end{defn}
that is, the set of configurations $\mathbf{c} \in
\mathcal{C}^*(S)$ not dominated by any other configuration.
Pareto-optimal configurations are configurations belonging to the
Pareto-optimal set and the {\em Pareto-optimal front} is the image
of the Pareto-optimal configurations, i.e. the set:
\[ \mathcal{P}_{F}(S,b) = \{ \mathbf{o} : \mathbf{o} = E(\mathbf{c},b), \mathbf{c} \in \mathcal{P}(S,b) \} \]

@@

\begin{figure}[h]
\includegraphics[width=0.9\columnwidth]{img/Pareto_set_and_front}
\caption{Relation between Pareto set (its ``parameter space representation'')
and Pareto front (``objective space representation'')}
\end{figure}

\subsection{Sketch of the Idea}

@@

We also consider, in measuring how much iteresting a region is, the
``innovation'' that it adds. If, during a design space exploration,
a Pareto front is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable ``innovation'',
because the way it fulfils the objectives is very similar to other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is very distant from previous ones is remarkable,
because it let the experimenter realize that the objectives can be
fulfilled in a completely different way from what he already found
during exploration. So a region will be evaluated ineresting as much
as how different are its added Pareto front points from the ones already
found.

%TODO: sync name of values with formulation
\subsection{Required Definitions}
\begin{defn}
\emph{Parameter interval}

@@

\begin{rem}
We will consider only parameters $p_{i}$ with ordering, i.e. such
that, if $P_{i}$ is a domain, $\forall a,b\in P_{i}$ it is possible
to say $a<b$ or $a=b$ or $a>b$.
\end{rem}
%TODO: DESCRIVERE COME SI PENSA DI TRATTARE GLI SCENARI IN CUI ESISTONO
%PARAMETRI SENZA ORDINAMENTO{]}
\begin{defn}
\emph{\label{pers02.def:Contiguous-intervals}Contiguous intervals}

@@

\comment{ CERCARE DI FAR CAPIRE CONCETTUALMENTE IL SEGUENTE CONFRONTO CON LA SENSITIVITY ANALYSIS:
It is worth noticing that this approach can not capture
``local sensitivity'' (this drawback will be called ``local sensitivity
limitation''), i.e. the objectives may be more sensitive to some
ranges of a parameter and less sensitive to other ranges of the same
parameter. Moreover it can not capture ``combined sensitivity''
(``combined sensitivity limitation''), i.e. the objectives may be
more sensitive to a range of a parameter $p_{1}$, only when other
parameters are within certain ranges, and less sensitive to the same
range of parameter $p_{1}$ when the other parameters are not within
those ranges. Therefore, in the method we propose, we do not use the
sensitivity directly, but we use the concept of innovation (that can
be considered, at least weakly related to the sensitivity). We do
not measure innovation of a parameter (as done for the sensitivity),
but rather we measure the innovation of a ``region'' of parameter
space, in order to overcome both the ``local sensitivity limitation''
and the ``combined sensitivity limitation'' stated above for the
sensitivity analysis.}

### 2014-07-23 introduction.tex
AA: intro + related work reviewed

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases an high level
system simulation is often the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, easily resulting in an untractable number
of simulations. In these cases a ``Design Space
Exploration'' strategy is required, i.e. a methodology to evaluate
only a small subset of all possible configurations while maintaining a
good level of efficiency and accuracy~\cite{surviving_soc}.

The final result of a design space exploration is a subset of the
configurations called Pareto Set (\cite{pareto}) (See also Definition
\ref{pers02.def:Pareto-set}. It represents the trade-offs between
objectives, so that for each of configuration of the Pareto Set there is no
other configuration performing better for all the objectives
considered. Once these few promising configurations have been obtained, a
subsequent step of the design flow can eventually afford a more
accurate low level simulation.

In this work we present PS, a multiobjective design exploration
strategy that introduces the concept of Parameter Space Representation
of Pareto Set. The main idea is focusing on the ``innovation''
introduced in the Pareto sets and then accordingly distribuiting the
exploration effort in regions created in the multidimensional space of
system parameters.

Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

Some exploration strategies assume some kind of ``knowledge'' about
the system parameters, their meaning and impact on design objectives.
The ``Dependency analysis'' proposed in \cite{givargis_tvlsi02}, is meant
to take advantage of the dependency among the parameters in a multiobjective
optimization problem.  With this a-priori knowledge, the designer can construct a
``dependency graph'' and recognize clusters, i.e. subsets of strongly
dependency-connected parameters. Each cluster is exhaustively evaluated
and its ``local Pareto set'' is found. Then, all local Pareto sets
are merged. Doing this way, a ``sum'' of local exhaustive evaluations
are performed instead of an exhaustive evaluation of all the possible
configurations. Some problems arise: i) In real scenarios, it may
be very difficult to recognize really independent clusters of parameters,
because there may be interdependencies among a very large number of
parameters. This may lead to the need of an exhaustive search through
almost all the possible configurations. ii) A designer may not have
a precise and complete picture of the dependencies among parameters;
for this reason the need of ``automated approaches for computing interdependencies'' is declared in the same paper. 
These drawbacks are not present in our contribution. The algorithm presented
here is also meant to take advantage of dependency but is not directly
based on dependency itself, but rather it ``catches'' dependencies
in terms of ``interesting or uninteresting regions''; as a consequence
no a-priori knowledge of dependencies is required as they are discovered
during exploration. Moreover, recognizing regions offers the capability
to capture local dependencies,  i.e. dependencies that
emerge only if parameter values are bounded to certain ranges.

Abraham, Rau and Schreiber proposed in \cite{santosh_hptr00} to decompose
the system to be evaluated into components that interact minimally
with each other. Pareto sets for each component are found and, provided
that ``monotonicity'' exists, the complete Pareto set is computed
merging the component Pareto sets. Roughly speaking (see section 4
of the same paper for more details), monotonicity property guarantees
that the best system can be obtained as a composition of the best
components. This approach would perfectly work if all the components
were perfectly isolated, i.e., if there were no dependencies among
different components (in a sense similar (but not necesarily equal)
to the one of Dependency analisys). But real scenarios seldom if ever
expose monotonicity property. This leads to some ``inaccuracies''
(as stated in section 4.6.1 of the same paper).

Other approaches take into account the concept of ``sensitivity''.
The sensitivity of an objective with the respect to a parameter is
a measure of how much the objective varies when varying that parameter.
A ``sensitivity analysis'' was used in \cite{fornaciari_codes01} in an optimization
problem with only one objective (the power-delay product of an electronic
device). This method was extended in \cite{palesi_iwsoc02} to a multiobjective
optimization. In that work the sensitivity of objectives with
the respect to each parameter is measured, and parameters are ordered
on the base on these sensitivity measures. Then, a first Pareto set
is obtained varying the first parameter, having fixed to arbitrary
values all other parameters. The Pareto set is then refined varying,
in a similar way, the other parameters, one by one. Experimental results
(see section 4.2 of the same work) show that this method is not of
good ``quality'' (as broadly defined in the introduction). The reason
can be found in the very limited and rigid exploration of the parameter
space: after computing the first Pareto set, there are no more chances
to consider configurations with values of the first parameter other
than the ones assumed in the first Pareto set calculation. Similar
limitations are imposed after computing the second Pareto set, the
third and so on. It is worth noticing that this approach can not capture
``local sensitivity'' (this drawback will be called ``local sensitivity
limitation''), i.e. the objectives may be more sensitive to some
ranges of a parameter and less sensitive to other ranges of the same
parameter. Moreover it can not capture ``combined sensitivity''
(``combined sensitivity limitation''), i.e. the objectives may be
more sensitive to a range of a parameter $p_{1}$, only when other
parameters are within certain ranges, and less sensitive to the same
range of parameter $p_{1}$ when the other parameters are not within
those ranges. Therefore, in the method we propose, we do not use the
sensitivity directly, but we use the concept of innovation (that can
be considered, at least weakly related to the sensitivity). We do
not measure innovation of a parameter (as done for the sensitivity),
but rather we measure the innovation of a ``region'' of parameter
space, in order to overcome both the ``local sensitivity limitation''
and the ``combined sensitivity limitation'' stated above for the
sensitivity analysis.

Many studies focus on genetic approaches to resolve multiobjective
optimization problems (\cite{coello_easmop} and others). Genetic approaches
have many advantages: they have proved good quality (as broadly defined
in the introduction), they are customizable, they are very general
and require no a-priori knowledge to the designer.  The strong
point of genetic approaches can be summarized saying that they consist
in a broad exploration (the mutation permits to make the exploration
not rigidly restricted to limited parts of parameter space), although
this exploration is not random but is guided by the performances of
the already evaluated configurations. Therefore, the exploration is
neither too much constrained nor too much random. We claim that the
approach presented in this paper has most of the benefits of the genetic
approaches, although the rational is completely different.

>>>>

The design of an embedded system requires different conflicting
objectives (energy consumption, performance, area) to be fulfilled.
Several factors are involved in this goal and real cases show how
often no analytical results are known to predict the relation between
system configurations and objectives. In such cases a high level
system simulation is often the only way to have a picture of how system
parameters impact on the objectives.  A key problem is that the size
of the design space to be simulated grows with the product of
cardinalities of each parameter, easily resulting in an untractable number
of simulations. In these cases a ``Design Space
Exploration'' strategy is required, i.e. a methodology to evaluate
only a small subset of all possible configurations while maintaining a
good level of efficiency and accuracy~\cite{surviving_soc}.

The final result of a design space exploration is a subset of the
configurations called Pareto Set (\cite{pareto}) (See also Definition
\ref{pers02.def:Pareto-set}). It represents the trade-offs between
objectives, so that for each configuration of the Pareto Set there is no
other configuration performing better for all the objectives
considered. Once these few promising configurations have been obtained, a
subsequent step of the design flow can eventually afford a more
accurate low level simulation.

In this work we present PS, a multiobjective design exploration
strategy that introduces the concept of Parameter Space Representation
of Pareto Set. The main idea is focusing on the ``innovation''
introduced in the Pareto sets and then accordingly distributing the
exploration effort in regions created in the multidimensional space of
system parameters. \comment{Potrebbe essere pompato di piu' con qualche cosa piu' d'impatto? }

\section{Related work}
\comment{Preferisci che i related work siano in un'unica sezione con Intro?}
Many different design space exploration algorithms have been proposed
in literature.  The motivations of an exploration
algorithm are rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

Some exploration strategies assume some kind of knowledge about
the system parameters, their meaning and their impact on design objectives.
The \emph{Dependency analysis}, proposed in \cite{givargis_tvlsi02}, is meant
to take advantage of the parameter dependencies, a-priori known by the designer \comment{E' giusto, cosi'?}. Starting from them, he can construct a
``dependency graph'' and recognize clusters, i.e. subsets of strongly
dependency-connected parameters. Each cluster is exhaustively evaluated
and its ``local'' Pareto set is found. Then, all local Pareto sets
are merged. In this way, a series of local exhaustive evaluations
are performed instead of an exhaustive evaluation of all the possible
configurations. Some problems arise: i) In real scenarios, it may
be difficult to recognize really independent clusters of parameters,
because there may be interdependencies among a large number of
parameters. This may lead to the need of an evaluation of
almost all the possible configurations. ii) A designer may not have
a precise and complete a-priori picture of the dependencies among parameters;
for this reason the need of ``automated approaches for computing interdependencies'' is declared in the same paper. 
These drawbacks are not present in our algorithm, since, although it leverages dependencies as in \cite{givargis_tvlsi02}, it is able to detect them automatically, with no a-priori knowledge required. Moreover, our algorithm is able to capture also ``local dependencies", i.e. dependencies that
emerge only among certain ranges of parameters and may not hold when considering the entires ranges. This cannot be modeled in Dependency Analysis.

Abraham, Rau and Schreiber proposed in \cite{santosh_hptr00} to decompose
the system under evaluation into components that interact minimally
with each other. Pareto sets for each component are found and, provided
that ``monotonicity'' exists, the complete Pareto set is computed
merging the component Pareto sets. Roughly speaking (see section 4
of the same paper for more details), monotonicity property guarantees
that the best system can be obtained as a composition of the best
components. This approach would perfectly work if all the components
were perfectly isolated, i.e., if there were no dependencies among
them. But real scenarios seldom if ever
expose monotonicity property thus resulting in inaccuracies, as stated in the same paper).

Other approaches, as \cite{fornaciari_codes01,palesi_iwsoc02}, are based the concept of \emph{sensitivity analysis}, i.e. measuring of how much the objective varies when varying each parameter.
Parameters are ordered based on their sensitivity. The first parameter (the most sensitive one) is varied, while keeping the other parameters fixed to arbitrary values, and its best value is found. The limited accuracy shown by these approaches can be explained by the limited and rigid exploration of the parameter
space: after fixing the best first parameter value, there are no more chances to consider different values. It is worth noticing that this approach can not capture
``local sensitivity'', i.e. the objectives may be more sensitive to some
ranges of a parameter and less sensitive to other ranges of the same
parameter. Moreover it can not capture ``combined sensitivity'', i.e. the objectives may be
more sensitive to a range of a parameter $p_{1}$, only when other
parameters are within certain ranges, and less sensitive to the same
range of $p_{1}$ when the other parameters are in different ranges. Otherwise, our algorithm can both capture local and combined sensitivity.

Many studies, as \cite{coello_easmop} and others, focus on genetic approaches to solve multiobjective
optimization problems. Genetic approaches
have many advantages: they provide good accuracy, they are customizable and very general
and require no a-priori knowledge to the designer.  The strong
point of genetic approaches can be summarized saying that they consist
in exploration that is sufficiently broad (the mutation avoids being rigidly restricted to limited parts of parameter space), and, at the same time, not too scattered as it is guided by the performances of
the already evaluated configurations. We claim that the
approach presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.

### 2014-07-23 paper.tex
AA: intro + related work reviewed

@@

\newcommand{\etal}{\emph{et al.}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\etc}{\emph{etc.}}
%\renewcommand{\baselinestretch}{0.98}\normalsize
%------------------------------------------------------------------------------
\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{definition}
\newtheorem{defn}{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{rem}{\protect\remarkname}

@@

\maketitle
\begin{abstract}
Embedded Systems Design often requires conflicting objectives to be
optimised with an approapriate choice of hardware/software parameters
. Interdependencies among these parameters can make this task even
more difficult, since an a-priori knowledge of them would  be required
to develop an analytical model of the system. In this work we present
PS, a multiobjective exploration strategy that introduces the notion
of Parameters Space Representation. After a detailed formal
description of the algorithm and its concepts, we show an
hardware/software exploration case study. Qualitative and quantitative
comparisons of PS against a well-know multiobjective genetic approach
demonstrate that, while not outperforming it in terms of dominance, the
proposed approach can trade the uniformity and granularity
qualities of the solutions found for obtaining larger Pareto sets,
thus representing a further choice for designer when objective constrains allow it.
\end{abstract}

>>>>

\usepackage{color}

@@

\newcommand{\etal}{\emph{et al.}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\etc}{\emph{etc.}}

\newcommand{\comment}[1]{\textcolor{red}{AA: \bf #1}}
%\newcommand\comment[1]{} % When you enable this lines, all the notes will disappear

%\renewcommand{\baselinestretch}{0.98}\normalsize
%------------------------------------------------------------------------------
\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{definition}
\newtheorem{defn}{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{rem}{\protect\remarkname}

@@

\maketitle
\begin{abstract}
Embedded Systems Design often requires conflicting objectives to be
optimized with an appropriate choice of hardware/software parameters
. Interdependencies among these parameters can make this task even
more difficult, since an a-priori knowledge of them would  be required
to develop an analytical model of the system. In this work we present
PS, a multiobjective exploration strategy that introduces the notion
of Parameters Space Representation. After a detailed formal
description of the algorithm and its concepts, we show a
hardware/software exploration case study. Qualitative and quantitative
comparisons of PS against a well-know multiobjective genetic approach
demonstrate that, while not outperforming it in terms of dominance, the
proposed approach can trade the uniformity and granularity
qualities of the solutions found for obtaining larger Pareto sets,
thus representing a further choice for designer when objective constrains allow it.
\end{abstract}
